---
title: "1_DiD_modeling: Big Data coursework paper"
author: "Ileana Marroquin"
date: "2023-11-15"
output:
 html_document: 
   toc: yes
   toc_depth: 2
   toc_float: yes
   theme: paper
   highlight: tango
   css: my.css
editor_options: 
  markdown: 
    wrap: 72
---

**READ ME**

**Topic:** "Applying Difference-in-difference to the World Bank Climate
aid allocation in Latin America. What has changed after the 2018 Escazu
Agreement?"

**Study Hypothesis:**

-   Ho: There is no significant difference in the amount of climate
    finance allocated by the WBG in the LAC countries based in their
    Escazu Agreement Signatory Status (SS) after 2018.

**Input:** the "classified_projects.csv" processed database

**Output:** DID model tables, validity test table and placebo test that
are relevant to answer the hypothesis

------------------------------------------------------------------------

# 0. Settup

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

```{r load-packages, message=FALSE, warning=FALSE}
rm(list=ls())

# Loading required libraries 
library(tidyverse)
library(here)
library(stringr)
library(dplyr)
library(texreg)
library(plm)
library(modelsummary)
library(broom)
library(knitr)
```

```{r load-database}
idf <- read_csv(
                here("1_processed_data","classified_projects.csv"),
                na = "")
glimpse(idf) #It is needed to transform this df to long format
```

# 1. Data base Modeling preparation and sub-setting

```{r reshaping}
# Preparing the data - from wide to long
idfw <- idf %>% 
  ungroup() %>%
  mutate(green = ifelse(environmental == 1, p.invest, NA),
         other = ifelse(environmental == 0, p.invest, NA)) %>%
  select(-p.invest) %>%
  group_by(project.id, p.year, country) %>%
  summarise(green = sum(green, na.rm = TRUE),
            other = sum(other, na.rm = TRUE)) %>%
  ungroup()

# Adding total year investment
idfw$wbinv <- idfw$green + idfw$other

idfw$wbinv <- log(idfw$wbinv)
idfw [ is.na(idfw) | idfw=="-Inf" ] = NA

glimpse(idfw)
```

# 2. Modeling: OLS and dynamic Diff-in-Diff

### 2.0 Final Construction Tasks for Analysis

```{r country-scoring}
# Introducing dummy for signture status (ss), 21 yes and 12 no
score <- structure(list(word = c("antigua and barbuda",
                                 "argentina",
                                 "bolivia",
                                 "brazil",
                                 "chile",
                                 "colombia",
                                 "costa rica",
                                 "dominica",
                                 "dominican republic",
                                 "ecuador",
                                 "guyana",
                                 "haiti",
                                 "honduras",
                                 "mexico",
                                 "nicaragua",
                                 "panama",
                                 "peru",
                                 "st kitts and nevis",
                                 "st lucia",
                                 "st vincent and the grenadines",
                                 "uruguay")))
str(score)

idfw$country <- trimws(idfw$country)
idfw$ss<- ifelse(idfw$country %in% score$word , 1, 0)
str(idfw)
```

### Observational Data Exploring

*Observational model simple difference in means after the change 2018
Escazu agreement treated group: 21 Signatory countries (SS=1) and
control group: Non-signatories (SS=0). Also note that columns of before
and after treatment need to be added*

```{r final-cleaning}
# Removing noisy regional projects
idfw <- subset(idfw, !(country %in% 
                         c("latin america",
                            "central america",
                            "caribbean",
                            "organization of eastern caribbean states")))
```

```{r subseting}
# Subseting years 2000 to 2033 - Future investment is also taken into account
idfw  <- idfw  %>%
                  rename( year = p.year)

did_subset <- idfw[idfw$year >= 2010 & idfw$year <= 2033, ]
```

```{r exploratory-data}

subset_mean <- did_subset %>%
                            group_by(year) %>%
                            summarise_at(vars(green, other),
                                         funs(mean(., na.rm=TRUE)))
subset_mean

# Evaluating the paralell trend assuption for DiD model
subset_mean %>%
            gather(variable, value, -year, na.rm = TRUE) %>%
            ggplot(aes(x = year, y = log(value), color = variable)) +
            geom_point() +
            geom_line() +
            ylab("Log of Investment in USD$") +
            xlab("Year") +
            guides(colour = guide_legend(title = "Type of Investment"))+
            ggtitle("WBG Projects Investment from 2010 to 2023")
```

## 2.1 Multivariate Regressions

### 2.1.1 OLS - Pre vs post

```{r OLS}
#Applying log 
did_subset$greenlog <- log(did_subset$green)
did_subset[is.na(did_subset) | did_subset=="-Inf"] = NA
did_subset$otherlog <- log(did_subset$other)
did_subset[is.na(did_subset) | did_subset=="-Inf"] = NA

#OLS regression signing country explaining type of investment
green_inv_reg <- lm(greenlog ~ ss, data = did_subset)
other_inv_reg <- lm(otherlog ~ ss, data = did_subset)
summary(green_inv_reg)
summary(other_inv_reg)

screenreg(list(green_inv_reg, other_inv_reg), caption = "OLS simple difference",
          caption.above = TRUE)

```

### 2.1.2 OLS - Simple diff

```{r modeling-subseting}
# Subseting the data for pre and post period
pre_escazu <- subset(did_subset, year < 2018)
post_escazu <- subset(did_subset, year >= 2018)
```

```{r manual-diff}
# Runing the simple difference model for pre and post period separately
# pre period
pre_escazu_reg = lm(wbinv ~ ss, data = pre_escazu)
summary(pre_escazu_reg) 

# post period
post_escazu_reg = lm(wbinv ~ ss, data = post_escazu)
summary(post_escazu_reg)
```

```{r summary of pre and post investment}
# Exporting the output table
screenreg(list(pre_escazu_reg, post_escazu_reg), 
          caption = "Pre and Post simple difference 2000-2033",
          caption.above = TRUE)
```

```{r regression of treatment and control groups}
# Pooled OLS Regression - Overall Investment = SS
t_invest_reg <- lm(wbinv ~ ss, data = did_subset)
summary(t_invest_reg)
```

## 2.2 ESCAZU 2018 DID MODEL

### 2.2.1 Diff-in-Diff modeling

```{r duplicates-test}
# Identify duplicates
library(dplyr)

duplicates <- did_subset %>%
              group_by(project.id, year) %>%
              count() %>%
              filter(n > 1)

# View the duplicates
print(duplicates)
```

```{r did-regression}
#DID model for ESCAZU 2018

# Creating a dummy for post escazu (treatment)
post_escazu <- ifelse(did_subset$year >= 2018, 1, 0)

# Subseting the data using the post-escazu dummy
escazu_did <- did_subset[did_subset$year < 2018 | did_subset$year >= 2018, ]

# Running the DiD model
escazu_model <- plm(wbinv ~ ss * post_escazu, data = escazu_did, 
                 index = c("year", "project.id"), method = "within")
msummary(escazu_model, stars = c('*' = .1, '**' = .05, '***' = .01))
```

### 2.2.2 Extended Analysis: Dynamic Diff-in-Diff

|                |
|:---------------|
| To be released |

# 3. Validity Test and Robbustness checks

## 3.1 Haussman Test

```{r mean-calculations, echo=TRUE}
did_subset_validity <- did_subset %>%
  mutate(mean_wbinv = mean(wbinv, na.rm = TRUE),
         mean_ss = mean(ss, na.rm = TRUE))

fe_model <- plm(wbinv - mean_wbinv ~ ss - mean_ss + factor(year),
               data = did_subset_validity,
               index = c("country", "year"),
               model = "within")
```

```{r random-effects}
re_model <- plm(wbinv - mean_wbinv ~ ss + factor(year),
                data = did_subset_validity,
                index = c("country", "year"),
                model = "random")
summary(re_model, stars = c('*' = .1, '**' = .05, '***' = .01))
```

```{r haussmann-results}
# Validity Summary
hausmann_test <- phtest(fe_model, re_model)
hausmann_test
```

## 3.2 Placebo test: PARIS 2015 MODEL

```{r paris-subset}
# Subset the data for pre and post period
pre_paris <- subset(did_subset, year < 2015)
post_paris <- subset(did_subset, year >= 2015)

## Run the simple difference model for pre and post period separately

## pre period
pre_paris_reg = lm(wbinv ~ ss, data = pre_paris)
summary(pre_paris_reg) 

## post period
post_paris_reg = lm(wbinv ~ ss, data = post_paris)
summary(post_paris_reg)
```

```{r placebo-did}
#DID model for PARIS 2015
# Creating a dummy for post escazu (treatment)
post_paris <- ifelse(did_subset$year >= 2015, 1, 0)

# Subseting the data using the post-escazu dummy
did_data_p <- did_subset[did_subset$year < 2015 | did_subset$year >= 2015, ]

# Running the DiD model
did_model_paris <- plm(wbinv ~ ss * post_paris, data = did_data_p, 
                 index = c("year", "project.id"), method = "within")
msummary(did_model_paris, stars = c('*' = .1, '**' = .05, '***' = .01))
```
